---
title: "3_data_visualization"
author: "Colby S."
date: "4/12/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
### Exercises 3.2

1.  Run `ggplot(data = mpg)`.
    What do you see?
```{r}
ggplot(data = mpg)
```
    
2.  How many rows are in `mpg`?
    How many columns?
```{r}
nrow(mpg)
ncol(mpg)
```
    
3.  What does the `drv` variable describe?
    Read the help for `?mpg` to find out.
```{r}
names(mpg)
?mpg
"drv is he type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd"
```

4.  Make a scatterplot of `hwy` vs `cyl`.
```{r}
ggplot(data = mpg) +
  geom_point(aes(x = hwy, y = cyl))
```

5.  What happens if you make a scatterplot of `class` vs `drv`?
    Why is the plot not useful?
```{r}
ggplot(data = mpg) +
  geom_point(aes(x = class, y = drv))
"the plot is not useful because it doent communicate an effective story"
```
    

### Exercises

1.  What's gone wrong with this code?
    Why are the points not blue?

    ```{r}
    #| fig-alt: "Scatterplot of highway fuel efficiency versus engine size of cars in ggplot2::mpg that shows a negative association. All points are red and the legend shows a red point that is mapped to the word 'blue'."
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy, color = "blue")) +
      labs(
        caption = "Scatterplot of highway fuel efficiency versus engine size of cars in ggplot2::mpg that shows a negative association. All points are red and the legend shows a red point that is mapped to the word 'blue'."
      )
    
    "the color = 'blue' is declared in the aesthetics of the graph so the program is treating 'blue' as a varaible instead of the color. The default color is red.
    ```

2.  Which variables in `mpg` are categorical?
    Which variables are continuous?
    (Hint: type `?mpg` to read the documentation for the dataset).
    How can you see this information when you run `mpg`?
```{r}
?mpg
meta_mpg <- data.frame(
  var = names(mpg),
  category = c(T, T, F, F, T, T, T, F, F, T, T)
)
meta_mpg$category <- sub(T, "categorical", meta_mpg$category)
meta_mpg$category <- sub(F, "continuous", meta_mpg$category)

meta_mpg
```
    

3.  Map a continuous variable to `color`, `size`, and `shape`.
    How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cty))
cat("continous variable cannot be mapped into shape")

```

4.  What happens if you map the same variable to multiple aesthetics?
```{r}
cat("there is no error. Somtimes this is useful if one of the aestetics is fallable, like if someone is colorblind they can still tell by shape")
```

5.  What does the `stroke` aesthetic do?
    What shapes does it work with?
    (Hint: use `?geom_point`)
```{r}
"stroke affects the size of the point"
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), stroke = 1)
?geom_point
```


6.  What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?
    Note, you'll also need to specify x and y.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cty, color = displ <5))

```

### Exercises

1.  What happens if you facet on a continuous variable?
```{r}
"it looks bad or you get an error"
```

2.  What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean?
    How do they relate to this plot?
```{r}
"emty cells indicate there were no cars with that combination of drive train and cylinders"
```


3.  What plots does the following code make?
    What does `.` do?

    ```{r}
    #| eval: false
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(drv ~ .)
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(. ~ cyl)
    
    ". ignores that dimension"
    ```

4.  Take the first faceted plot in this section:

    ```{r}
    #| eval: false
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_wrap(~ class, nrow = 2)
    ```

    What are the advantages to using faceting instead of the colour aesthetic?
    What are the disadvantages?
    How might the balance change if you had a larger dataset?

5.  Read `?facet_wrap`.
    What does `nrow` do?
    What does `ncol` do?
    What other options control the layout of the individual panels?
    Why doesn't `facet_grid()` have `nrow` and `ncol` arguments?

6.  Which of the following two plots makes it easier to compare engine size (`displ`) across cars with different drive trains?
    What does this say about when to place a faceting variable across rows or columns?

    ```{r}
    #| fig-alt: "Two faceted plots, both visualizing highway fuel efficiency versus engine size of cars in ggplot2::mpg, faceted by drive train. In the top plot, facet are organized across rows and in the second, across columns."
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_grid(drv ~ .)
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_grid(. ~ drv)
    ```

7.  Recreate this plot using `facet_wrap()` instead of `facet_grid()`.
    How do the positions of the facet labels change?

    ```{r}
    #| fig-alt: "Scatterplot of highway fuel efficiency versus engine size of cars in ggplot2::mpg, faceted by type of drive train across rows."
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_wrap(drv ~ .)
    ```
    
### Exercises

1.  What geom would you use to draw a line chart?
    A boxplot?
    A histogram?
    An area chart?

2.  Run this code in your head and predict what the output will look like.
    Then, run the code in R and check your predictions.

    ```{r}
    #| eval: false
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ```

3.  What does `show.legend = FALSE` do?
    What happens if you remove it?\
    Why do you think I used it earlier in the chapter?

4.  What does the `se` argument to `geom_smooth()` do?

5.  Will these two graphs look different?
    Why/why not?

    ```{r}
    #| eval: false
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth()
    ggplot() + 
      geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
    ```

6.  Recreate the R code necessary to generate the following graphs.
    Note that wherever a categorical variable is used in the plot, it's `drv`.

    ```{r}
    #| echo: false
    #| message: false
    #| fig-width: 3
    #| out-width: "50%"
    #| fig-align: "default"
    #| fig-alt: "There are six scatterplots in this figure, arranged in a 3x2 grid. In all plots highway fuel efficiency of cars in ggplot2::mpg are on the y-axis and engine size is on the x-axis. The first plot shows all points in black with a smooth curve overlaid on them. In the second plot points are also all black, with separate smooth curves overlaid for each level of drive train. On the third plot, points and the smooth curves are represented in different colours for each level of drive train. In the fourth plot the points are represented in different colours for each level of drive train but there is only a single smooth line fitted to the whole data. In the fifth plot, points are represented in different colours for each level of drive train, and a separate smooth curve with different line types are fitted to each level of drive train. And finally in the sixth plot points are represented in different colours for each level of drive train and they have a thick white border."
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(aes(group = drv), se = FALSE) +
      geom_point()
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) +
      geom_smooth(aes(linetype = drv), se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(size = 4, colour = "white") + 
      geom_point(aes(colour = drv))
    ```
    
### Exercises

1.  What is the default geom associated with `stat_summary()`?
    How could you rewrite the previous plot to use that geom function instead of the stat function?

2.  What does `geom_col()` do?
    How is it different to `geom_bar()`?

3.  Most geoms and stats come in pairs that are almost always used in concert.
    Read through the documentation and make a list of all the pairs.
    What do they have in common?

4.  What variables does `stat_smooth()` compute?
    What parameters control its behaviour?

5.  In our proportion bar chart, we need to set `group = 1`.
    Why?
    In other words what is the problem with these two graphs?
    
### Exercises

1.  What is the problem with this plot?
    How could you improve it?

    ```{r}
    #| fig-alt: "Scatterplot of highway fuel efficiency versus city fuel efficiency of cars in ggplot2::mpg that shows a positive association. The number of points visible in this plot is less than the number of points in the dataset."
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
      geom_point()
    ```

2.  What parameters to `geom_jitter()` control the amount of jittering?

3.  Compare and contrast `geom_jitter()` with `geom_count()`.

4.  What's the default position adjustment for `geom_boxplot()`?
    Create a visualization of the `mpg` dataset that demonstrates it.
    
### Exercises

1.  Turn a stacked bar chart into a pie chart using `coord_polar()`.

2.  What does `labs()` do?
    Read the documentation.

3.  What's the difference between `coord_quickmap()` and `coord_map()`?

4.  What does the plot below tell you about the relationship between city and highway mpg?
    Why is `coord_fixed()` important?
    What does `geom_abline()` do?

    ```{r}
    #| fig-asp: 1
    #| out-width: "50%"
    #| fig-alt: "Scatterplot of highway fuel efficiency versus engine size of cars in ggplot2::mpg that shows a negative association. The plot also has a straight line that follows the trend of the relationship between the variables but doesn't go through the cloud of points, it's beneath it."
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() + 
      geom_abline() +
      coord_fixed()
    ```

