---
title: "ch_15_factors"
author: "Colby S."
date: "5/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

data(package = 'forcats')
```


```{r}
month.name
months <- c("April", "July", "January", "December", "July")
levels(months)
months2 <- factor(months, levels= month.name)
sort(months2)



count(gss_cat[[race]])

gss_cat %>%
  count(race)
```




Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?
```{r}
unique(gss_cat$rincome)

ggplot(data = gss_cat) +
  geom_bar(mapping =aes(x = rincome)) +
  coord_flip()



gss_cat %>%
  filter(!rincome %in% c("Not applicable")) %>%
  mutate(rincome = fct_recode(rincome,
    "Less than $1000" = "Lt $1000"
  )) %>%
  mutate(rincome_na = rincome %in% c("Refused", "Don't know", "No answer")) %>%
  
  ggplot(aes(x = rincome, fill = rincome_na)) +
  scale_fill_manual(values = c("FALSE" = "black", "TRUE" = "gray")) +
  geom_bar() +
  coord_flip() +
  scale_y_continuous("Number of Respondents", labels = scales::comma) +
  scale_x_discrete("Respondent's Income") +
  theme(legend.position = "None")
```

What is the most common relig in this survey? What’s the most common partyid?
```{r}
gss_cat %>%
  count(c(relig)) %>%
  arrange(desc(n)) %>%
  head(1)

gss_cat %>%
  count(c(partyid)) %>%
  arrange(desc(n)) %>%
  head(1)
```

Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?
```{r}
names(gss_cat)

gss_cat %>%
  group_by(relig) %>%
  distinct(denom) %>%
  arrange(desc(relig))

gss_cat %>%
  ggplot() +
  geom_bar(aes(x=denom)) +
  facet_wrap(~relig) +
  coord_flip()

```

### 15.4 Exercises
There are some suspiciously high numbers in tvhours. Is the mean a good summary?
```{r}
gss_cat %>%
  ggplot() +
  geom_bar(aes(tvhours))

# calc the mean tv hours watched per day, the mean age, and the number of observations FOR EACH RELIGION
gss_cat %>%
  group_by(relig) %>%
  summarise(
    tv_mean = mean(tvhours, na.rm = TRUE),
    age_mean = mean(age, na.rm = TRUE),
    n = n() 
  )
```

For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.
```{r}
# keep function to tell if variables are factors
keep(gss_cat, is.factor) %>%
  names()
```

Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?
```{r}
"the factors in first position are closest to the bottom"

```

### 15.5 Exercises 
How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?
```{r}
gss_cat %>%
  mutate(partyid = fct_collapse(
    partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  count(year, partyid) %>%
  
  group_by(year) %>%
  mutate(p = n / sum(n)) %>%
  
  ggplot() +
  geom_line(aes(
    x = year, y = p, 
    color = fct_reorder2(partyid, year, p))) +
  
  scale_fill_manual(values = c("ind" = "gray",
                               "dem" = "blue",
                               "rep" = "red",
                               "other" = "gray")) + #why doesnt this work
  labs(color = "Party")
  
         
```

How could you collapse rincome into a small set of categories?

```{r}
unique(gss_cat$rincome)

gss_cat %>%
  mutate(rincome = fct_collapse(rincome,
               "low"= c(str_c("$", c(1000, 3000, 4000, 5000), " to ", c(2999, 3999, 4999, 5999))))) %>%
  ggplot() +
  geom_bar(aes(x = rincome)) +
  coord
```

