---
title: "ch_13_exercises"
author: "Colby S."
date: "4/27/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("nycflights13")
library("viridis")
# library("datamodelr") #not available in this version of r


airlines <- nycflights13::airlines
planes <- nycflights13::planes
weather <- nycflights13::weather
airports <- nycflights13::airports
flights <- nycflights13::flights
```
### 13.2.1 Exercises
Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

the locations of the origin and destination
```{r}
data(package = "nycflights13")
airlines <- nycflights13::airlines
planes <- nycflights13::planes
weather <- nycflights13::weather
airports <- nycflights13::airports # airports contains lat long of origin and destination airports
flights <- nycflights13::flights

flights %>%
  select(origin, dest) %>%
  inner_join(select(airports, origin = "faa", origin_lat = "lat", origin_lon = "lon", origin_name = "name"), by = "origin") %>%
  inner_join(select(airports, dest = "faa", dest_lat = "lat", dest_lon = "lon", , dest_name = "name"), by = "dest") %>%
  
  slice(1:100) %>%
  
  ggplot(aes(x = origin_lon, xend = dest_lon,
             y = origin_lat, yend = dest_lat)) +
  borders("state") +
  geom_segment(arrow = arrow(length = unit(0.2, "cm"))) +
  coord_quickmap() +
  labs(x = "Longitude", y = "Latitude")

```


I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?
```{r}
weather
airports
 "origin and faa are keys to one another"
```

weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?


### 13.3.1 Exercises
Add a surrogate key to flights.
```{r}
flights %>%
  mutate(surrogate_key = row_number()) %>%
  glimpse()
```

Identify the keys in the following datasets

```{r}
# Lahman::Batting
names(Lahman::Batting)
Lahman::Batting %>%
  count(playerID, yearID, stint) %>%
  filter(n > 1) %>%
  nrow()

# babynames::babynames
names(babynames::babynames)
babynames::babynames %>%
  count(year, sex, name) %>%
  filter(n > 1) %>%
  nrow()

# nasaweather::atmos
names(nasaweather::atmos)
nasaweather::atmos %>%
  count(lat, long, year, month) %>% 
  filter(n > 1) %>%
  nrow()

nasaweather::atmos %>%
  summarise(n_distinct(lat, long, year, month) == nrow(nasaweather::atmos))

# fueleconomy::vehicles
names(fueleconomy::vehicles)
fueleconomy::vehicles %>%
    summarise(n_distinct(id) == nrow(fueleconomy::vehicles))

# ggplot2::diamonds
names(ggplot2::diamonds)
ggplot2::diamonds %>%
  summarise(nrow(distinct(.)) > nrow(ggplot2::diamonds)) #there are duplicate rows in the data, a surrogate key would be needed
```

Draw a diagram illustrating the connections between the Batting, People, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between People, Managers, AwardsManagers.
```{r}
sum(names(Lahman::Batting) %in% names(Lahman::People))
glimpse(c(Lahman::Batting, Lahman::People))
Lahman::People
Lahman::Salaries
```

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

```{r}

```






### 13.4.6 Exercises
Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}

flights %>%
  filter(!is.na(arr_delay)) %>%
  group_by(dest) %>%
  summarise("mean_delay" = mean(arr_delay)) %>%
  inner_join(airports, by = c(dest = "faa")) %>%
  ggplot() +
  borders("state") +
  geom_point(aes(x = lon, y = lat, size = mean_delay, alpha = mean_delay)) +
  coord_quickmap()
  
  
```

airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.

Add the location of the origin and destination (i.e. the lat and lon) to flights.
```{r}
airports
flights %>%
  inner_join(select(airports, origin = "faa", origin_lon = "lon", origin_lat = "lat") , by = "origin" ) %>%
  inner_join(select(airports, dest = "faa", dest_lon = "lon", dest_lat = "lat") , by = "dest" )
  
  
```

Is there a relationship between the age of a plane and its delays?
```{r}
glimpse(planes)
glimpse(flights)


planes %>%
  summarise(n_distinct(tailnum) == nrow(planes)) # tailnum is primary key for planes df 

flights %>%
  inner_join(select(planes, tailnum, plane_made = "year"), by = "tailnum") %>%
  mutate(age = year - plane_made) %>%
  select(age, dep_delay, arr_delay) %>%
  
  pivot_longer(cols = 2:3, 
               names_to = "type", 
               values_to = "delay") %>%
  
  mutate(age = if_else(age > 25, 25L, age)) %>%
  filter(!is.na(age), !is.na(delay)) %>%
  group_by(age, type) %>%
  summarise(
    mean_delay = mean(delay)
  ) %>%
  ggplot() +
  geom_point(aes(x = age, y = mean_delay, color = type))
  

  # view()
  
  ggplot() +
  geom_boxplot(aes(x = as.factor(age), y = delay))
  
```

What weather conditions make it more likely to see a delay?
```{r}

```

What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.
```{r}

```

### 13.5.1 Exercises
What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

Filter flights to only show flights with planes that have flown at least 100 flights.

Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?

You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

